RNeXML: Parsing and Serializing the Next Generation of Phyloinformatic Data in R

### 2014-03-12 inst/doc/pubs/manuscript.Rmd
updated manuscript, add test for new bug

Though R provides the most extensive and widely used collection of comparative phylogenetic methods tools, R users lack a convenient format to exchange and archive comparative methods data.

@@

A nexml file can be thought of as having three components: (1) phylogeny data (2) character-trait data, and (3) metadata. `RNeXML` seeks to provide a seemless way to convert these to and from their corresponding R object representations, such as `ape::phylo` phylogenies, character matrices (`data.frames`) and metadata lists (or for advanced users, `XMLInternalDocument` and `rrdf` triplestores).  The RNeXML package functions (the package Application Programming Interface, API)

@@

<!-- Do we describe installation of the package?  `install.packages("RNeXML")`?-->

@@

We begin by reading in an example NeXML file.

@@

[@ouch]: http://cran.r-project.org/web/packages/ouch/
[@phylobase]: http://cran.r-project.org/web/packages/phylobase

@@

@@

[Temple Lang 2013, XML]: http://cran.r-project.org/web/packages/XML/

@@

[Wickham 2011]: http://vita.had.co.nz/papers/testthat.html

@@

>>>>

Though R provides the most extensive and widely used collection of comparative phylogenetic methods tools, R users lack a convenient, extensible format to exchange and archive comparative methods data.

@@

@@

A nexml file can be thought of as having three components: (1) phylogeny data (2) character-trait data, and (3) metadata. `RNeXML` seeks to provide a seemless way to convert these to and from their corresponding R object representations, such as `ape::phylo` phylogenies, character matrices (`data.frames`) and metadata lists (or for advanced users, `XMLInternalDocument` and `rrdf` triplestores). Metadata can provide information about the study in which the phylogeny was generated or applied (such as authors, citations to publications, timestamps, and even descriptions of methods), but can also be used to annotate individual data components.  We illustrate how metadata annotations can be added to the operational taxonomic units of the phylogeny and can be used to extend the data representation to new uses such as stochastic character maps.

The current stable version of the `RNeXML` package can be installed from the CRAN library <!-- Actually not yet -->
using the standard

```r
install.packages("RNeXML")
```

The latest stable development version can be installed from Github using the `devtools` package:

```r
library(devtools)
install_github("RNeXML", "ropensci")
```

@@

We begin by reading in an example NeXML file provided with the package.

@@

@@

This allows us to include multiple trees and character matrices to a single NeXML file.

@@

@@

@@

[Temple Lang 2013, XML]: http://cran.r-project.org/web/packages/XML/
[Wickham 2011]: http://vita.had.co.nz/papers/testthat.html
[Chamberlain 2013]: http://dx.doi.org/10.12688/f1000research.2-191.v2
[@ouch]: http://cran.r-project.org/web/packages/ouch/
[@phylobase]: http://cran.r-project.org/web/packages/phylobase
